---
title: "Reproducible Research: Peer Assessment 1, by Tony"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity_data<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
mean_step_by_day<-aggregate(steps~date,data=activity_data,mean)
median_step_by_day<-aggregate(steps~date,data=activity_data,median)
sum_step_by_day<-aggregate(steps~date,data=activity_data,sum)
hist(sum_step_by_day[,2],
     main="Histogram of Mean steps by day",
     xlab="total steps by day",breaks=20)
```

Below are the mean and median steps per day.
```{r}
mean(sum_step_by_day[,2])
median(sum_step_by_day[,2])
```

## What is the average daily activity pattern?
```{r}
# Subset by intervals by 5
mean_step_by_day_5<-aggregate(steps~date,
      data=activity_data[activity_data$interval==5,],
      function(x) mean(x,na.rm=TRUE))
```
Plot time series
```{r}
plot(as.Date(mean_step_by_day_5$date),mean_step_by_day_5$steps,type="l",
     xlab="Date",ylab="steps")
```

The date with maximum steps on interval 5 is:
```{r}
mean_step_by_day_5[mean_step_by_day_5$step==
    max(mean_step_by_day_5$step),]
```



```
## Imputing missing values

The number of NA values are:
```{r}
sum(is.na(activity_data$steps))
```

We will put the means of the all the steps in the NAs
```{r}
mean_of_all_steps <- mean(activity_data$steps,na.rm=TRUE)
NA_filled_activity_data <- activity_data
NA_filled_activity_data$steps[is.na(NA_filled_activity_data$steps)] <-
  mean_of_all_steps


mean_step_by_day2<-aggregate(steps~date,data=NA_filled_activity_data,mean)
median_step_by_day2<-aggregate(steps~date,data=NA_filled_activity_data,median)

sum_step_by_day2<-aggregate(steps~date,data=NA_filled_activity_data,sum)
hist(sum_step_by_day2[,2],
     main="Histogram of Mean steps by day",
     xlab="total steps by day",breaks=20)

```
Below are the mean and median steps per day with NAs changed to means.
```{r}
mean(sum_step_by_day2[,2])
median(sum_step_by_day2[,2])
```
As one can see, changing the NAs to the mean of the steps per day had no change to the mean and little change to the median


## Are there differences in activity patterns between weekdays and weekends?
```{r}

# Subset by intervals by 5
NA_mean_step_by_day_5<-aggregate(steps~date,
      data=NA_filled_activity_data[NA_filled_activity_data$interval==5,],mean)
#Make a column with day of week
NA_mean_step_by_day_5$day_of_week<-
  weekdays(as.Date(NA_mean_step_by_day_5$date))
#Make Saturdays and Sundays to "Weekend"
NA_mean_step_by_day_5$day_of_week[
    NA_mean_step_by_day_5$day_of_week== "Saturday" |
    NA_mean_step_by_day_5$day_of_week== "Sunday"]<-
  "Weekend"
#Make non-Weekend to "Weekday"
NA_mean_step_by_day_5$day_of_week[
    NA_mean_step_by_day_5$day_of_week != "Weekend"]<-
  "Weekday"
```
```{r}
#Make it into a factor
NA_mean_step_by_day_5$day_of_week<-as.factor(NA_mean_step_by_day_5$day_of_week)

#Now plot using gglot
library(ggplot2)
library(plyr)
# Change date into formal date that can be plotted
NA_mean_step_by_day_5 <- mutate(NA_mean_step_by_day_5, date = as.Date(NA_mean_step_by_day_5$date))
ggplot(data = NA_mean_step_by_day_5, aes(date, steps)) + 
    geom_line() +facet_grid(day_of_week~.)

```


